Conversations in Time: Interactive Visualization to Explore Structured Temporal Data

### 2021-07-29 rjdistill_version/tsibbletalk.Rmd
more fixes about rjdistill

Figure \@ref(fig:highlight-retail) illustrates examining temporal data with many keys. The data has 152 series corresponding to different industries in retail data. The multiple series are displayed using an overlaid time series plot, along with a scatterplot of two variables (trend versus seasonal strength) from feature space, where each series is represented by a dot. The feature space is computed using the `features()` function from \CRANpkg{feasts}, which summarises the original data for each series using various statistical features. This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \@ref(fig:highlight-retail-2).

```{r highlight-retail, fig.height = 6, out.width = "49%", fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a).", fig.subcap = c("", ""), fig.ncol = 2}
```

@@

The tree plot provides the graphics skeleton, upon which the rest of the data plots can be attached. In this example, small multiples of line plots are placed at the top right of Figure \@ref(fig:tourism-linking-fig) to explore the temporal trend across regions by the trip purpose. The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.

@@

There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \@ref(fig:tourism-linking-fig).  (This is the interactive realization of Figure \@ref(fig:highlight-retail)). <!-- Routine functions bring about new interaction with temporal data on the client side.-->

@@

The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture the daily rhythms of the downtown [@ped]. Figure \@ref(fig:wrap-ped-html) shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months. Plot \@ref(fig:wrap-ped-1) shows hourly counts from January to May on an absolute timeline, facetted by locations. The stage 3 COVID-19 lockdown, on March 16, is marked by a change of color. (The pre-lockdown period is colored with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations. QV Market is less affected probably because this is a major produce market, an essential service that continued to operate. Bourke St, a major shopping center, sees a gradual uptick in the last weeks of the period indicating that people were getting back into the shops.

@@

The wrapping procedure involves slicing the time index into seasonal periods of interest, and the result is diced time. For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \@ref(fig:wrap-ped-2). For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly. This can be achieved using shiny, with the functions provided in the \CRANpkg{tsibbletalk}.

This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable component that could be embedded in any shiny application. In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers. As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied id string that must be unique. The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly. The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine, so that the time `index` can be retrieved for dissection. The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc. In other words, the `period` defines the grind level. For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year". The following code snippet generates Figure \@ref(fig:wrap-ped). The creation of the `pedestrian20` data is available in supplementary R files.

@@

Figure \@ref(fig:wrap-ped-1) corresponds to the initial state, with the slider incremented by 1-day units. The "play" button near the end of the slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days. Alternatively, users can drag the slider to examine selected slices.

>>>>

Figure \@ref(fig:highlight-retail) illustrates examining temporal data with many keys. The data has 152 series corresponding to different industries in retail data. The multiple series are displayed using an overlaid time series plot, along with a scatterplot of two variables (trend versus seasonal strength) from feature space, where each series is represented by a dot. The feature space is computed using the `features()` function from \CRANpkg{feasts}, which summarises the original data for each series using various statistical features. This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, which can be graphically displayed as in Figure \@ref(fig:highlight-retail) (right).

```{r highlight-retail, fig.height = 3.6, fig.show = "hold", fig.cap = "Plots for the \\code{aus\\_retail} data, with the series of strongest seasonal strength highlighted. (a) An overlaid time series plot. (b) A scatter plot drawn from their time series features, where each dot represents a time series from (a)."}
```

@@

The tree plot provides the graphics skeleton, upon which the rest of the data plots can be attached. In this example, small multiples of line plots are placed at the top right of Figure \@ref(fig:tourism-multi) to explore the temporal trend across regions by the trip purpose. The shared tsibble data can be directly piped into \CRANpkg{ggplot2} code to create this.

@@

There is one final step, to compose the three plots into an ensemble of coordinated views for exploration, shown in Figure \@ref(fig:tourism-multi).  (This is the interactive realization of Figure \@ref(fig:highlight-retail)). <!-- Routine functions bring about new interaction with temporal data on the client side.-->

@@

The city of Melbourne has sensors installed at various locations, to record hourly counts of pedestrians, in order to capture the daily rhythms of the downtown [@ped]. Figure \@ref(fig:wrap-ped-html) shows the first five months of 2020 foot traffic at four different locations, for three different time slices, daily, weekly and full five months. Plot \@ref(fig:wrap-ped-html) shows hourly counts from January to May on an absolute timeline, facetted by locations. The stage 3 COVID-19 lockdown, on March 16, is marked by a change of color. (The pre-lockdown period is colored with dark green and lockdown with orange.) We can see a significant decline in foot traffic  at all four locations. QV Market is less affected probably because this is a major produce market, an essential service that continued to operate. Bourke St, a major shopping center, sees a gradual uptick in the last weeks of the period indicating that people were getting back into the shops.

@@

The wrapping procedure involves slicing the time index into seasonal periods of interest, and the result is diced time. For example, hourly pedestrian data can be decomposed into 24-hour blocks, which then overlays the counts for all respective days, as done in plot \@ref(fig:wrap-ped-html). For exploration, this slice position should be controlled interactively, so that many different slices can be examined rapidly. This can be achieved using shiny, with the functions provided in the \CRANpkg{tsibbletalk}.

This shiny module, decoupled to `tsibbleWrapUI()` and `tsibbleWrapServer()`, presents a clean interface and forms a reusable component that could be embedded in any shiny application. In general, a shiny module provides a vehicle for modularising shiny applications, relevant for both users and developers. As with all shiny modules, the first argument in both functions in \CRANpkg{tsibbletalk} requires a user-supplied id string that must be unique. The UI function `tsibbleWrapUI()` simply shows a slider that animates or controls the number of periods to be diced. The workhorse is the server function `tsibbleWrapServer()`, encapsulating the algorithm that transforms data and sends messages to update the plot accordingly. The `plot` argument expects a `ggplot` or `plotly` object, where one can plot data using either lines or other graphical elements (such as boxplots). As the function name suggests, a (shared) tsibble is needed to start the engine, so that the time `index` can be retrieved for dissection. The `period` option semantically takes a desired number of seasonal periods to be shifted, for example data shifted by "1 day", "2 days", or "1 week", etc. In other words, the `period` defines the grind level. For date-times (represented by `POSIXt`), the granularity ranges from fine "day" to a much coarser "year". The following code snippet generates Figure \@ref(fig:wrap-ped-html). The creation of the `pedestrian20` data is available in supplementary R files.

@@

Figure \@ref(fig:wrap-ped-html) corresponds to the initial state, with the slider incremented by 1-day units. The "play" button near the end of the slider can automatically animate the slicing and dicing process, walking the viewer through all 24 hours of the 152 days. Alternatively, users can drag the slider to examine selected slices.
